Article Title

### 2014-01-29 plos.tex
Rev44: submitted version

\paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
The estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.  These could have multiple physiological causes such as common inputs, cell assembly activation, network fluctuations, or other forms of synchronized activations.

@@

In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}). The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons. The low-rank component had rank 15, expressing 15 inferred latent units. The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{06}{A and B}). There was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions. This variability may be explained by differences in brain states and recording quality and needs further investigation.

@@

While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\%) of negative values (\figref{05}{F}). This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.  This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions have been uncovered.

@@

Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.  We therefore compared the interactions inferred by $C_{\sf sparse+latent}$ to those obtained by thresholding the sample correlations with the threshold set to match the connectivity of the $C_{\sf sparse+latent}$ graph.  The networks revealed by the two methods differed substantially. In the example site with 17.8\% connectivity inferred by $C_{\sf sparse+latent}$, only 41.7\% of connections coincided with above-threshold sample correlations (\figref{05}{F, H, and I}). In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.  This low overlap between the two types of graphs was consistent across the datasets \figref{06}{E}. 
t

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites} Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size in each imaged site.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size at each imaged site.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percentage overlap between the highest sample correlations and the  interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity.
}
\label{fig:06}}
{\includegraphics{./figures/Figure06.pdf}}
\end{figure}

@@

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.  Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.

@@

In our study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits. We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.  The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution.  Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations. %\Kcomment{I am not sure about the previous argument.}

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies. The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}. However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

Several strategies have been developed to limit the model space in order to improve the quality of the estimate. For example, \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina. Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, with limited data an incorrect, but simpler (lower dimension) model can produce a better estimate than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

@@

Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit. Importantly, the inferred functional interactions were substantially different from those measured by considering the network of the most significant sample correlations \figref{05}{F, H, and I}.  For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) that may reflect inhibitory circuitry.  Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles. Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry. Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation. For example, these latent units may be modulated under different brain states (e.g. slow-wave sleep, attention) and stimulus conditions (e.g. certain types of stimuli may engage feedback connections).

@@

Calcium imaging began 1 hour after dye injection.  All imaging was performed using the 3D-RAMP two-photon microscope as described in \cite{Cotton:2013}. First, a 3D stack was acquired and cells manually segmented. Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.

\paragraph{Visual stimulus}
Full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s. Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations. Directional tuning was map using a pseudo-random sequence of drifting gratings at sixteen equally spaced directional of motion changing at 2 Hz for 3 min without blanks. The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.  Each stimulus condition was presented at least 180 times. 
\paragraph{Data processing}
All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt datajoint.github.com}) first developed in our lab.

@@

Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.  The $p$-value of the fit was computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.  Cells were considered tuned for $p$-values below 0.05.

For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).  Spike trains that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site), presumably from glial cells or inactive neurons, were exculded from the analysis.

\paragraph{Cross-validation}
To compare the performance of the estimator against each other, we used conventional 10-fold cross validation. Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This splitting of data was chosen to achieve sufficient independence between the 10 datasets for cross-validaton purposes. Splitting the data into smaller blocks would increase their dependence due to temporal effects. Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}.  The outer loop evaluated the performance of the estimators with optimal values of the hyperparameters.  The optimization of the hyperparameters was performed within the inner loop in two phases: random search to find a good starting point and pattern search to find the global minimum.  The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.  Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.

When cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.  This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
\paragraph{Covariance Estimation}
Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.  The computation of regularized estimators only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.

Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.  The variances (the diagonal of $C_{\sf 0}$) were shrunk toward (linearly mixed with) their mean value:
\begin{equation}
D = (1-\alpha)C_{\sf 0}\circ I + \alpha \frac 1 p \Tr(C_{\sf 0}) I
\end{equation}
Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.

The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the normal loss function (Eq.~\ref{eq:loss}).

Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
The factor matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
\begin{equation}
(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L + \tilde\Psi,C_{\sf 0}},
\end{equation}
which we solved by an expectation-maximization (EM) algorithm.  Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.  The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).

Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$. The precision matrix is computed in two steps: First, the zero structure is determined by minimizing the $L_1$-penalized loss:
\begin{equation}
Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf 0}} + \lambda \|\tilde Z \|_1
\end{equation}
where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$. This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}. To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}. 
Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
\begin{equation}
S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf 0}},
\end{equation}
where $Z^\sharp$ denotes the set of positive-definite matrices with zeros in all entries where entries of $Z$ equal zero.  This step was also solved by ADMM.  Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}). Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine tune the regularization level.

@@

The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component: 
\begin{equation}
P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

@@

\paragraph{Simulation}
For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units. The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}. Samples were then generated by multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.

>>>>

\paragraph{Structure of $C_{\sf sparse+latent}$ estimates}
The estimator $C_{\sf sparse+latent}$ separates two sources of correlations: a network of linear interactions between pairs of neurons and a set of latent units reflecting common fluctuations across the entire population.

@@

In the example site (Fig.~\ref{fig:05}), the sparse component had 82.2\% sparsity (or conversely, 17.8\% connectivity: $connectivity=1-sparsity$) with average node degree of 52.9 (\figref{05}{G}). The averge node degree, \ie the average number of interactions linking each neuron, is related to connectivity as $degree = connectivity\cdot(p-1)$, where $p$ is the number of neurons. The low-rank component had rank 15, denoting 15 inferred latent units. The number of latent units and the average node degrees generally increased with population size, although these values varied widely between the datasets (\figref{06}{A and B}). There was an inverse relationship between the number of latent units and the average node degree (\figref{06}{C}): Several sites, despite their large population sizes, were driven by latent units and had few pairwise interactions. This variability may be explained by differences in brain states and recording quality and needs further investigation.

@@

While the sample correlations were mostly positive, the sparse component of the partial correlations (\sq{interactions}) had a high fraction (43.3\% in the example site) of negative values (\figref{05}{F}). This fraction was consistent across multiple sites but the fraction of negative interactions increased with the inferred connectivity \figref{06}{F}.  This increase suggests that negative interactions are inferred only after a sufficient density of positive interactions has been uncovered.

@@

Thresholded sample correlations have been used in several studies to infer pairwise interactions \cite{Golshani:2009, Feldt:2011, Malmersjo:2013}.  We therefore compared the interactions inferred by $C_{\sf sparse+latent}$ to those obtained by thresholding the sample correlations with the threshold set to match the connectivity of the $C_{\sf sparse+latent}$ graph.  The networks revealed by the two methods differed substantially. In the example site with 17.8\% connectivity inferred by $C_{\sf sparse+latent}$, only 41.7\% of connections coincided with above-threshold sample correlations (\figref{05}{F, H, and I}). In particular, most of the inferred negative interactions corresponded to low sample correlations (\figref{05}{F}) where high correlations were expected given the rest of the correlation matrix.  This low overlap between the two types of graphs was consistent across the datasets (\figref{06}{E}).

\begin{figure}    \floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,center},capbesidewidth=8.3cm}}]{figure}[\FBwidth]
    {\caption{{\bf Properties of $C_{\sf sparse+latent}$ estimates from all imaged sites.} 
    Each point represents an imaged site with its color indicating the population size as shown in panels A and B. The example site from Figures \ref{fig:03} and \ref{fig:05} is circled in blue. 
\\
    {\bf A.} The number of inferred latent units vs.~population size.
    {\bf B.} The average node degree for sparse partial correlations vs.~population size.
    {\bf C.} The average node degree in the sparse component in each site vs.~the number of latent units described by the low-rank component of the estimate.
    {\bf D.} The average sample correlations  vs.~the average partial correlations (Eq.~\ref{eq:partial}) of the $C_{\sf sparse+latent}$ estimate.
    {\bf E.} Percent overlap between the graph of highest sample correlations and the graph of interactions inferred by $C_{\sf sparse+latent}$ as a function of the connectivity (\ie $1-sparsity$) of its sparse component. 
    {\bf F.} The percentage of negative interactions vs.~connectivity in the $C_{\sf sparse+latent}$ estimates.
}
\label{fig:06}}
{\includegraphics{./figures/Figure06.pdf}}
\end{figure}

@@

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models constrained by observed pairwise correlations also model pairwise coupling between cells \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.  Assuming that the population response follows a multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to anatomy.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question and we did not attempt them here.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe the functional connectivity in cortical populations until this study.

@@

In our study, we sought to infer pairwise functional connectivity networks  in cortical microcircuits. We hypothesized that partial correlations correspond more closely to underlying mechanisms than sample correlations when recordings are sufficiently dense.  Since neurons form synaptic connections mostly locally and sparsely \cite{Perin:2011}, we \emph{a priori} favored solutions with sparse partial correlations.  The importance of partial correlations may be justified by the principle of maximum entropy: The maximum entropy distribution on discrete or continuous multivariate domains constrained by the observed mean firing rates, their variances, and correlations is the multivariate normal distribution, where the precision matrix specifies the interactions between cell pairs.  Therefore, under the aforementioned assumptions that the recorded population is sufficiently complete and that the model correctly represents the nature of interactions, the network of partial correlations can be hypothesized to be a better representation of functional dependencies than correlations. %\Kcomment{I am not sure about the previous argument.}

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies. The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As techniques have improved to allow recording from larger neuronal populations, experimental neuroscientists have addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}. However, ambitious projects, such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed by using regularized estimators. Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hinders interpretation.

Several strategies have been developed to limit the model space in order to improve the quality of the estimate. For example, Ganmor et al. \cite{Ganmor:2011} developed a heuristic rule to identify the most significant features that must be fitted by a maximum entropy model for improved performance in retina. Generalized linear models typically employ $L_1$ penalty terms to constrain the solution space and to effectively reduce the dimensionality of the solution \cite{Pillow:2008}.

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, with limited data, an incorrect but more constrained model can produce a better estimate than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not necessarily support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary low-dimensional target is guaranteed to produce \emph{some} improvement over the best estimate with many degrees of freedom.

@@

Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit. Importantly, the inferred functional interactions were substantially different from the network of the most significant sample correlations \figref{05}{F, H, and I}.  For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) and may reflect inhibitory circuitry.  Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity, it may accurately reflect average connectivity profiles. Future experiments utilizing molecular markers for different cell types and follow-up multiple whole-cell \emph{in vitro} experiments \cite{Hofer:2011, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry. Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation. For example, these latent units may be modulated under different brain states (e.g. slow-wave sleep, attention) and stimulus conditions (e.g. certain types of stimuli may engage feedback connections).

@@

Calcium imaging began 1 hour after dye injection.  All imaging was performed using the 3D-RAMP two-photon microscope as described in \cite{Cotton:2013}. First, a 3D stack was acquired and cells were manually segmented. Then calcium signal were collected by sampling in the center of each cell at rates of 100 Hz or higher, depending on the number of cells.

\paragraph{Visual stimulus}
The visual stimulus consisted of full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s. Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations. Directional tuning was mapped using a pseudo-random sequence of drifting gratings at sixteen equally spaced directions of motion, 500 ms per direction, for 3 min without blanks. The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 datasets) or five directions (in 22 datasets) were used and the presentations lasted 1 second and were separated by 1-second blanks.  Each stimulus condition was presented between 100 and 300 times.
\paragraph{Data processing}
All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt datajoint.github.com}) first developed in our lab.

@@

Orientation tuning was computed by fitting the mean firing rates in response to gratings drifting in directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[\frac 1 w(\cos(\phi-\theta)-1) \right] + c\exp\left[\frac 1 w(\cos(\phi-\theta+\pi)-1) \right]$ where $b\ge c$ are the amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.  The $p$-value of the fit was computed as the fraction of the permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.  Cells were considered tuned with $p$-values below 0.05.

For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 14 and 27 minutes (mean 22 minutes).  Cells that did not have substantial spiking activity (those whose variance was less than 2\% of the $90^{th}$ percentile in each site) were exculded from the analysis.

\paragraph{Cross-validation}
To compare the performance of the estimators against each other, we used conventional 10-fold cross validation. Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This splitting of the data was chosen to achieve a high degree of independence between the 10 datasets for cross-validaton purposes. Splitting the data into smaller blocks could increase dependencies between them due to temporal correlations thereby violating the assumptions of cross-validation. Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

Since each of the regularized estimators had one or two hyperparameters, we used \emph{nested cross-validation}:  The outer loop evaluated the performance of the estimators with the optimal hyperparameter values estimated within the inner loop.  Hyperparameters were optimized by a two-phase search algorithm: random search to find a good starting point followed by pattern search to find the global minimum.  The inner cross-validation loop subdivided the training dataset from the outer loop to perform 10-fold cross-validation in order to evaluate each choice of the hyperparameter values.  Thus the size of the training dataset within the inner loop comprised 81\% of the entire recording.

When the cross-validation loss was not required, only the inner loop of cross-validation was used, applied to the entire dataset.  This approach was used to compute the covariance matrix estimates and their excess-loss in the simulation study (\figref{02}{\,Rows 4 and 5}) and to analyze the partial correlation structure of the sparse+latent estimator (Fig.~\ref{fig:05}--\ref{fig:07}).
\paragraph{Covariance estimation}
Within the inner loop of cross-validation, covariance matrix estimation was performed with fixed hyperparameter values provided by the search algorithm.  The computation of regularized estimators only required the sample covariance matrix $C_{\sf 0}$ of the training dataset.

Estimator $C_{\sf diag}$ (Eq.~\ref{eq:c-diag})  used two hyperparameters: the covariance shrinkage intensity $\lambda \in [0,1]$ and variance shrinkage intensity $\alpha \in [0,1]$.  The variances (the diagonal of $C_{\sf 0}$) were shrunk toward (linearly mixed with) their mean value:
\begin{equation}
D = (1-\alpha)C_{\sf 0}\circ I + \alpha \frac 1 p \Tr(C_{\sf 0}) I
\end{equation}
Then the diagonal matrix $D$ was used as the target of covariance shrinkage (Eq.~\ref{eq:c-diag}) to produce the final regularized estimate.  The {\tt corpcor} package in the programming language R also implements this estimator \cite{Schafer:2010}, although its analytical optimization of the shrinkage intensities is based on the mean squared error whereas we optimized them with respect to the normal loss function (Eq.~\ref{eq:loss}).

Estimator $C_{\sf factor}$ used two hyperparameters: the number of latent factors $d$ and the shrinkage intensity $\lambda \in [0, 1]$.  
The matrix $L$ of rank $d$ and the diagonal matrix of individual variances $\Psi$ were computed by solving the minimization problem
\begin{equation}
(L,\Psi) = \argmin\limits_{\tilde L,\tilde\Psi} \loss{\tilde L + \tilde\Psi,C_{\sf 0}},
\end{equation}
which we solved by an expectation-maximization (EM) algorithm.  Under our chosen loss function (Eq.~\ref{eq:loss}), this is equivalent to maximum likelihood estimation of $L$ and $\Psi$ under the multivariate Gaussian distribution.  The final regularized estimate is obtained by linear shrinkage toward the factor model (Eq.~\ref{eq:c-factor}).

Estimator $C_{\sf sparse}$  has one hyperparameter $\lambda$ to regulate the sparsity of its precision matrix $S$. The precision matrix is computed in two steps: First, the zero structure is determined by minimizing the $L_1$-penalized loss:
\begin{equation}
Z = \argmin\limits_{\tilde Z \succ 0} \loss{{\tilde Z}^{-1},C_{\sf 0}} + \lambda \|\tilde Z \|_1
\end{equation}
where $\tilde Z\succ 0$ denotes the constraint that $\tilde Z$ be a positive definite matrix and $\|\tilde Z\|_1$ is the element-wise $L_1$ norm of the matrix $\tilde Z$. This problem formulation is known as \emph{graphical lasso} \cite{Meinshausen:2006, Friedman:2008}. To solve this minimization problem, we modified the alternative-direction method of multipliers (ADMM) algorithm \cite{Ma:2013}. 
Then, after the zero structure was determined, the remaining coefficients were fitted without penalty:
\begin{equation}
S = \argmin\limits_{\tilde S \in Z^\sharp} \loss{\tilde S,C_{\sf 0}},
\end{equation}
where $Z^\sharp$ denotes the set of positive-definite matrices with the same zero structure as $Z$.  This step was also solved by ADMM.  Then the final estimate is the inverse of $S$ (Eq.~\ref{eq:c-sparse}). Unlike $C_{\sf diag}$ and $C_{\sf factor}$, this estimator does not include linear shrinkage: the selection of the sparsity level provides sufficient flexibility to fine tune the regularization level.

@@

The partial correlation matrix with both sparse and low-rank components is computed from $C_{\sf sparse+latent}$ according to Eq.~\ref{eq:partial}; it includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses of average partial correlations (\figref{05}{B and E} and \figref{06}{D}, and \figref{07}{A--C}).  The sparse component $S$ was used separately in analyses of the connectivity graphs (\figref{05}{F, G, H} and \figref{06}{B, C, E, and F} and \figref{07}{D--F}); in this case, the partial correlations were computed by normalizing the sparse component alone without the low-rank component: 
\begin{equation}
P_{\sf sparse} = -(S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

@@

\paragraph{Simulation}
For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units. The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,Row 2}). Samples were then drawn from multivariate normal distributions with the respective true covariance matrices to be estimated by each of the estimators.
